Технология использования открытых данных Федеральной налоговой службы России для экономико-географического анализа

### 2024-04-20 legal-companies-geography/Paper.Rmd
Finish results, add conclusion, #1

Text text text.

@@

**KEYWORDS:** w1, w2

@@

@@

>>>>

This paper provides an exploratory regional analysis of legal business in Russia for 2016–21. The analysis is based mainly on the Federal Tax Service (FTS) open data on small and medium-sized businesses and on the number of employees of organisations, as well as on portions of official statistical information. The paper proposes a reproducible technique of working with FTS open data for regional analysis in service geography that was partially implemented in a specially designed Python command-line tool. The analysis includes questions about the count of legal firms and their employees by region, about the share of organisations in regional legal business, about the role of legal business in region's economy, and about the yearly change of the number of legal companies in regions. In addition, it briefly addresses an issue of probable connection between the location of multi-regional courts and the development of legal business. It is found out that legal business in Russia is highly concentrated in Moscow and Saint Petersburg that jointly account for 1/3 of legal firms and workers and are also at the top by the role of legal business in the region. However, there are other regions with relatively high development of legal business, that are located mainly on the European South of Russia, in the Volga-Urals and in Siberia. There is a group of regions with domination of organisations in legal business and the group of regions where individual entrepreneurs prevail over companies. District courts are usually located in the regions with high level of legal business development, but the mechanism of this co-location requires further investigation. In general, the number of legal firms in regions grown from 2016 to 2021, with the common inverted U-shape pattern probably explained by the coronavirus (covid-19) pandemic that negetively impacted the small and medium business and reverted the initial growth trend.

@@

**KEYWORDS:** regional analysis, geography of services, legal services, legal business, open data

@@

The changes in the number of legal firms by region are depicted in Fig. 4. It can be clearly seen that the count of such companies increased in the majority of regions, indicating the development of legal business. Few regions have lost companies, all of them are located in the European part of Russia, and their net change in the number of legal firms from 2016 to 2021 is relatively low (usually less than 10%) compared to the increase rates in other regions. Sometimes neither decline or growth was not clearly expressed (for example, Altai republic, Lipetsk and Tambov regions). The regions with the largest growth are some republics at Notrh Caucas, as well as Kalmyk republic and Nenetz territory. Their results are probably explained by the low base effect, because the absolute counts of legal firms in all there regions are below 100 (see Fig. 1). As for other regions with high growth, Leningrad and Moscow regions should be remarked, because their results may be partially explained by the relocation of companies from the cities (Saint Petersburg and Moscow) to the surrounding regions. However, Moscow and Saint Petersburg also gained legal firms with one of the top net changes. The observed patterns of changes indicated by lines of the map can be divided into several groups. The first is the stable yet saturating increase typical for Moscow and Saint Petersburg regions, and, less expresses, Khabarovsk territory, Rostov and Ryazan regions. The second is the inverted U-shape curve which is charasteristic of the majority of regions with overall growth in the number of legal firms. This pattern can be explained by the coronavirus (COVID-19) pandemic effect that has negatively impacted small and medium-sized business and reverted the previous trend of the growth. The “height” of this inverted U varies by region (the highest observed in Chukotka territory, Tyva and Adygeya republics). It can be hypothesised that the first group is in fact a subset of the second, but the intrinsic growth rate before the pandemic was higher, resulting in the slowing down of the growth rather than in the final decrease. The third pattern is typical for the regions with overall decrease in the number of legal firms, and it can be described as “initial short-term growth, then continuous decline”. The fourth group is a mixture of increase and decrease, for example Nenetz territory and Northern Osetia republic.

```{r fig-4-timeline-map, fig.cap="Fig. 4. The change in the number of legal firms by region in 2016–2021"}
timeline_map
```

@@

The conducted regional analysis discovered several spatial features of legal business in Russia. First, the regions are quite different with respect to the count of legal firms or their employees. Several regional clusters with relatively high concentration of legal business can be found out, for example, the European South, Volga-Ural and Siberian. Unsurprisingly, Moscow and Saint Petersburg cities are characterized by the top count of legal firms and workers, but the overall concentration is really huge, because they jointly account for approximately 1/3 of Russia's legal businesses and legal companies employees. Besides, it's again Moscow and Saint Petersburg that have the highest share of legal companies employees in region's labor force (above 0.4% that is 2 or 3 times higher than the average value for other regions). However, there are some other regions with relatively high role of legal business in regional economy: Sverdlosvk, Tyumen and Novosibirsk.

Regions can be roughly grouped into two categories: one with the legal organisations prevailing in the legal business, and the other with individual entrepreneurs prevailing over companies. The high role of companies is typical for Moscow, Saint Petersburg, some Volga and Siberian regions, as well as a small portion of central regions (Vologda and Ivanovo), while individuals are important in the European South and North, in the majority of Central European regions, in several regions of the Urals, and in the Far East.

The role of region in country's legal business is usually associated with the presence of one or many multi-regional courts in the administrative center of this region. Only two regions do not act as a host for any of such the courts but still have relatively high level of development of legal business: Volgograd region and Bashkir republic. This observation may be explained in two ways: either the courts act as a magnets for legal business, because courts are necessary for many legal activities, or the courts are placed in regions and cities with highly developer legal business. Further research may throw the light on the direction of this court-business relation.

From 2016 to 2021, the number of legal firms in Russia's regions mostly increased. The decrease is characteristic only of few regions. The top growth was observed in Moscow, Saint Petersburg, as well as in the surrounding regions. In addition, some regions (mainly at the European South) also shown very high growth, but probably due to the low base effect. The yearly patterns of change may be categorized into four groups, and the inverted U-shape change explained likely by the coronavirus (covid-19) pandemic is the most common among the regions.

The results of this explanatory study provide an overview of Russia's recent legal business geography and may be used as a support for data-driven business and governmental decisions. Several research hypothesis also may be conducted from this analysis, thus this study may be treated as a base for further academic development in the field of legal services geography in Russia. Apart from this, the reproducible data-driven technique of this research based entirely on open data and open source software tools may be used (with minor changes) for regional analysis and mapping of other narrow service sector areas.
